```{r setup,include=FALSE}
# set the knitr options ... for everyone!
# if you unset this, then vignette build bonks. oh, joy.
#opts_knit$set(progress=TRUE)
opts_knit$set(eval.after='fig.cap')
# for a package vignette, you do want to echo.
# opts_chunk$set(echo=FALSE,warning=FALSE,message=FALSE)
opts_chunk$set(warning=FALSE,message=FALSE)
#opts_chunk$set(results="asis")
opts_chunk$set(cache=TRUE,cache.path="cache/")

#opts_chunk$set(fig.path="figure/",dev=c("pdf","cairo_ps"))
opts_chunk$set(fig.path="github_extra/figure/",dev=c("png"))
opts_chunk$set(fig.width=7,fig.height=6,dpi=100,out.width='700px',out.height='600px')

# doing this means that png files are made of figures;
# the savings is small, and it looks like shit:
#opts_chunk$set(fig.path="figure/",dev=c("png","pdf","cairo_ps"))
#opts_chunk$set(fig.width=4,fig.height=4)
# for figures? this is sweave-specific?
#opts_knit$set(eps=TRUE)

# this would be for figures:
#opts_chunk$set(out.width='.8\\textwidth')
# for text wrapping:
options(width=96,digits=2)
opts_chunk$set(size="small")
opts_chunk$set(tidy=TRUE,tidy.opts=list(width.cutoff=50,keep.blank.line=TRUE))
library(ggplot2)
```

# IMDb Mirror

This is a 'mirror' of the IMDb database, stored in a mariadb
database. The data are downloaded from one of the IMDb FTP
mirrors, then processed to remove TV shows, porn, shorts, 
and 'insignificant' films (based on number of ratings).
The database and scraper are housed in `docker` containers
and orchestrated via `docker-compose` 

## Initialization

Prepare the `srv` directory where we will store the database
and the make state:

```{r prepare,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
mkdir -p ./srv/imdb/mariadb ./srv/imdb/state/make ./srv/imdb/rw 
chmod -R 777 ./srv/imdb
```

Then use `docker-compose` to initialize the database.

```{r init,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
docker-compose build
docker-compose up -d
sleep 2s && docker-compose ps
```

It may take a while to pull the image for `mariadb`. To check on the status
of your images, run

```{r dcps,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
docker-compose ps
```

You should see something like the following:
```
         Name                       Command              State             Ports          
-----------------------------------------------------------------------------------------
imdbmirror_imdb_1         /usr/bin/make noop             Exit 0                           
imdbmirror_mariastate_1   /bin/true                      Exit 0                           
imdbmirror_mysqldb_1      /docker-entrypoint.sh mysqld   Up       0.0.0.0:23306->3306/tcp 
imdbmirror_state_1        /bin/true                      Exit 0                           
```

## Download

The commands to download, transform, and store the IMDb data are
run via a `Makefile` within the `docker` image `imdb`. The main command
of this `docker` image is `make`, so y
be run via:

```{r imake1,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
alias imake='docker-compose run --rm imdb'
imake help
```

Creating a mirror works by the following steps:

1. Downloading files from the FTP mirror.
1. Processing them with perl to remove TV shows.
1. Stuffing them into a sqlite file with `imdb2sql.py`.
1. Removing porn, unpopular titles and shorts, as well as stranded actors and
	 directors.
1. Stuffing the remaining data into mariadb, interpreting some of the fields.

You can run them via:

```{r imake2,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
# download:
imake -j 4 downloaded
# process 
imake -j 6 procd
# raw sqlite
imake raw_sqlite
# remove porn and so on:
imake sqlite
# stuff into mariadb:
imake stuffed
```

**Note** This can take a good long while. Depending on your internet connection, the 
download might take 20 minutes or more. Processing into sqlite may take around 10 to 
15 minutes, and then post-processing and loading into mariadb may take an hour. Also,
intermediate files will be kept around, saved in the `./srv/imdb/state` directory.
Having intermediate files allows you to change the code and rerun without forcing a
new download.

Now connect to the database and poke around:

```{r mysqlcon,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
mysql -s --host=0.0.0.0 --port=23306 --user=moe --password=movies4me IMDB
```

## What is in the database?

First, let us check the available tables:

```{r avail_tables,cache=TRUE,eval=TRUE,tidy=FALSE}
library(RMySQL)
library(dplyr)
library(knitr)
imcon <- src_mysql(host='0.0.0.0',user='moe',password='movies4me',dbname='IMDB',port=23306)
capt <- dbGetQuery(imcon$con,'SET NAMES utf8')
show(src_tbls(imcon))
```

### Movie info

First, how many movies do we have:

```{r avail_movies,cache=TRUE,eval=TRUE,tidy=FALSE}
nmovies <- tbl(imcon,'title') %>%
	summarize(count=n()) 
print(nmovies)
```

Now some information about each of the tables:


```{r movie_tables,cache=TRUE,eval=TRUE,tidy=FALSE}
showtable <- function(tnam) {
  cat(sprintf('%8s:\n',tnam))
	tbl(imcon,tnam) %>%
		left_join(tbl(imcon,'title') %>% select(movie_id,title),by='movie_id') %>%
    head() %>%
    kable(caption=tnam) %>%
		show()
}

movie_tables <- src_tbls(imcon) 
movie_tables <- movie_tables[grepl('^movie_',movie_tables)]
capt <- lapply(movie_tables,showtable)

```

## Example usage

Here we use `dplyr` to connect to the database, and create a list of
the best 'underrated' movies of each year. The criteria are:

1. More than 500 ratings, but
1. fewer than 2000 ratings, and 
1. ranked via mean rating divided by standard deviation.


```{r sortem,cache=TRUE,eval=TRUE,tidy=FALSE}
library(RMySQL)
library(dplyr)
library(knitr)
imcon <- src_mysql(host='0.0.0.0',user='moe',password='movies4me',dbname='IMDB',port=23306)
dbGetQuery(imcon$con,'SET NAMES utf8')

tbl(imcon,'movie_votes') %>%
  filter(votes >= 500,votes <= 2000) %>% 
  inner_join(tbl(imcon,'movie_runtime') %>% 
    filter(runtime > 60) %>% 
    select(movie_id),by='movie_id') %>%
  mutate(vote_sr=vote_mean/vote_sd) %>%
  select(movie_id,vote_sr) %>%
  left_join(tbl(imcon,'title') %>% select(movie_id,title,production_year),by='movie_id') %>%
  arrange(desc(vote_sr)) %>%
  collect() %>% 
  group_by(production_year) %>%
  summarize(title=first(title),
    sr=max(vote_sr)) %>%
  ungroup() %>%
  arrange(production_year) %>%
  kable()
```



