```{r setup,include=FALSE}
# set the knitr options ... for everyone!
# if you unset this, then vignette build bonks. oh, joy.
#opts_knit$set(progress=TRUE)
opts_knit$set(eval.after='fig.cap')
# for a package vignette, you do want to echo.
# opts_chunk$set(echo=FALSE,warning=FALSE,message=FALSE)
opts_chunk$set(warning=FALSE,message=FALSE)
#opts_chunk$set(results="asis")
opts_chunk$set(cache=TRUE,cache.path="cache/")

#opts_chunk$set(fig.path="figure/",dev=c("pdf","cairo_ps"))
#opts_chunk$set(fig.path="github_extra/figure/",dev=c("png"))
opts_chunk$set(fig.path="figure/README_",dev=c("png"))
opts_chunk$set(fig.width=9,fig.height=6,dpi=100,out.width='900px',out.height='600px')

# doing this means that png files are made of figures;
# the savings is small, and it looks like shit:
#opts_chunk$set(fig.path="figure/",dev=c("png","pdf","cairo_ps"))
#opts_chunk$set(fig.width=4,fig.height=4)
# for figures? this is sweave-specific?
#opts_knit$set(eps=TRUE)

# this would be for figures:
#opts_chunk$set(out.width='.8\\textwidth')
# for text wrapping:
options(width=96,digits=2)
opts_chunk$set(size="small")
opts_chunk$set(tidy=TRUE,tidy.opts=list(width.cutoff=50,keep.blank.line=TRUE))
library(ggplot2)
```

# IMDb Mirror

This is a 'mirror' of the IMDb database, stored in a mariadb
database. The data are downloaded from one of the IMDb FTP
mirrors, then processed to remove TV shows, porn, shorts, 
and 'insignificant' films (based on number of ratings).
The database and scraper are housed in `docker` containers
and orchestrated via `docker-compose` 

## Initialization

Prepare the `srv` directory where we will store the database
and the make state:

```{r prepare,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
mkdir -p ./srv/imdb/mariadb ./srv/imdb/state/make ./srv/imdb/rw 
chmod -R 777 ./srv/imdb
```

Then use `docker-compose` to initialize the database.

```{r init,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
docker-compose build
docker-compose up -d
sleep 2s && docker-compose ps
```

It may take a while to pull the image for `mariadb`. To check on the status
of your images, run

```{r dcps,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
docker-compose ps
```

You should see something like the following:
```
         Name                       Command              State             Ports          
-----------------------------------------------------------------------------------------
imdbmirror_imdb_1         /usr/bin/make noop             Exit 0                           
imdbmirror_mariastate_1   /bin/true                      Exit 0                           
imdbmirror_mysqldb_1      /docker-entrypoint.sh mysqld   Up       0.0.0.0:23306->3306/tcp 
imdbmirror_state_1        /bin/true                      Exit 0                           
```

## Download

The commands to download, transform, and store the IMDb data are
run via a `Makefile` within the `docker` image `imdb`. The main command
of this `docker` image is `make`, so y
be run via:

```{r imake1,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
alias imake='docker-compose run --rm imdb'
imake help
```

Creating a mirror works by the following steps:

1. Downloading files from the FTP mirror.
1. Processing them with perl to remove TV shows.
1. Stuffing them into a sqlite file with `imdb2sql.py`.
1. Removing porn, unpopular titles and shorts, as well as stranded actors and
	 directors.
1. Stuffing the remaining data into mariadb, interpreting some of the fields.

You can run them via:

```{r imake2,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
# download:
imake -j 4 downloaded
# process 
imake -j 6 procd
# raw sqlite
imake raw_sqlite
# remove porn and so on:
imake sqlite
# stuff into mariadb:
imake stuffed
```

**Note** This can take a good long while. Depending on your internet connection, the 
download might take 20 minutes or more. Processing into sqlite may take around 10 to 
15 minutes, and then post-processing and loading into mariadb may take an hour. Also,
intermediate files will be kept around, saved in the `./srv/imdb/state` directory.
Having intermediate files allows you to change the code and rerun without forcing a
new download.

Now connect to the database and poke around:

```{r mysqlcon,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
mysql -s --host=0.0.0.0 --port=23306 --user=moe --password=movies4me IMDB
```

## Shiny frontend

The `docker-compose` command now also supports a shiny database front end. If you wish to 
turn it off, comment out the following stanza from `docker-compose.yml` (I do not know
of a better solution at the moment):

```{r docompose1,engine='bash',comment='bash',cache=TRUE,eval=FALSE}
# comment it out by putting '#' in front, like a nice yaml file.
shiny:
  extends:
    file: common.yml
    service: common
  build: ./shiny
  links: 
  - mysqldb:mysqldb
  ports:
  - 0.0.0.0:19191:5555
```

If you _do_ want to use the shiny frontend, once the image is running, go to
[localhost port 19191](http://localhost:19191), and poke around. You can
grep for title, cast members, directors, restrict by votes per year and
production year. When you select a row from the table in the `titles` tab,
details on that film will become available in the other tabs.

---------------------

## What is in the database?

First, let us check the available tables:

```{r avail_tables,cache=TRUE,eval=TRUE,tidy=FALSE}
library(RMySQL)
library(dplyr)
library(knitr)
imcon <- src_mysql(host='0.0.0.0',user='moe',password='movies4me',dbname='IMDB',port=23306)
capt <- dbGetQuery(imcon$con,'SET NAMES utf8')
show(src_tbls(imcon))
```

## Movie info

First, how many movies do we have:

```{r avail_movies,cache=TRUE,eval=TRUE,tidy=FALSE}
nmovies <- tbl(imcon,'title') %>%
	summarize(count=n()) 
print(nmovies)
```

Now some information about each of the tables:


```{r movie_tables,cache=TRUE,eval=TRUE,tidy=FALSE,results='asis'}
library(RMySQL)
library(dplyr)
library(knitr)
imcon <- src_mysql(host='0.0.0.0',user='moe',password='movies4me',dbname='IMDB',port=23306)
capt <- dbGetQuery(imcon$con,'SET NAMES utf8')
showtable <- function(tnam) {
  cat(sprintf('\n\n### %8s:\n',tnam))
	tbl(imcon,tnam) %>%
		left_join(tbl(imcon,'title') %>% select(movie_id,title),by='movie_id') %>%
    head() %>%
    kable(caption=tnam) %>%
		show()
}

movie_tables <- src_tbls(imcon) 
movie_tables <- movie_tables[grepl('^movie_',movie_tables)]
capt <- lapply(movie_tables,showtable)

```

-------------

## Cast by budget

Here we grab all movies with budget information (_n.b._ these are sometimes
bogus, because IMDb users can enter the information without much confirmation.)
then sum movie budgets for each actor and actress appearing in the movie,
then find the top actor and actress by birth year. Here we list the top
actor and actress by birth year, with total budget in billions, and the number
of films:


```{r starpower_I,cache=TRUE,eval=TRUE,tidy=FALSE,results='asis'}
library(RMySQL)
library(dplyr)
library(knitr)
imcon <- src_mysql(host='0.0.0.0',user='moe',password='movies4me',dbname='IMDB',port=23306)
capt <- dbGetQuery(imcon$con,'SET NAMES utf8')

# values over 300M are likely bogus
budgets <- tbl(imcon,'movie_budgets') %>%
	filter(units=='$') %>%
	select(movie_id,amount) %>%
	filter(amount < 3e8)

role_info <- tbl(imcon,'role_type') %>%
	filter(role %in% c('actor','actress')) %>%
	select(role_id)

sumbudget <- tbl(imcon,'cast_info') %>%
	inner_join(role_info,by='role_id') %>%
  select(movie_id,person_id,role_id) %>%
	inner_join(budgets,by='movie_id') %>%
	group_by(person_id) %>% 
	summarize(totroles=n(),
    sumamount=sum(amount)) %>%
  ungroup()

aggregated <- sumbudget %>%
	inner_join(tbl(imcon,'name') %>%
		select(person_id,name,dob,gender) %>%
    filter(!is.null(dob)),by='person_id') %>%
  collect() 

topcast <- aggregated %>%
	mutate(yob=as.numeric(gsub('^((19|20)\\d{2}).+','\\1',dob))) %>%
	filter(!is.na(yob)) %>%
  group_by(yob,gender) %>%
  arrange(desc(sumamount)) %>%
  summarize(totbudget=first(sumamount) / 1e9,
    topname=first(name),
    toproles=first(totroles)) %>%
  ungroup()


# I wish there were an easy way to do this via tidyr,
# and maybe there is...
topm <- topcast %>% 
  select(yob,topname,toproles,totbudget,gender) %>% 
	filter(gender=='m')
topf <- topcast %>% 
  select(yob,topname,toproles,totbudget,gender) %>% 
	filter(gender=='f')

topc <- inner_join(topm %>% select(-gender),topf %>% select(-gender),by='yob') %>%
	rename(actor_total_budget=totbudget.x,
    actor_name=topname.x,
    actor_roles=toproles.x,
    actress_total_budget=totbudget.y,
    actress_name=topname.y,
    actress_roles=toproles.y)

topc %>% 
	filter(yob > 1945,yob <= 1995) %>%
  kable()
```

### Try again

That kind of works. However, the `cast_info` table also contains a field
`nr_order` that lists the _order_ of an actor/actress in some films. 
(In some titles, the `nr_order` field is not populated, or lists
each cast member as 1.) Presumably lower order equates to higher billing
in the cast and perhaps higher salary. Here we proceed to estimate
total earned salary for actors and actresses by assuming salary is inversely
proportional to the `nr_order`, as follows:

1. For each film, fill out any missing `nr_order` with the largest known
value for that film.
1. Then substitute all missing `nr_order` values with a 1 (this only
occurs for films with no recorded values for `nr_order`).
1. Rescale the `nr_order` for each film so that they sum to 1.0 for
each film.
1. Multiply the rescaled `nr_order` by one-seventh the budget of the film.
1. Aggregate over actors and actresses, then group by birth year and
gender.

```{r starpower_II,cache=TRUE,eval=TRUE,tidy=FALSE}
library(RMySQL)
library(dplyr)
library(knitr)
imcon <- src_mysql(host='0.0.0.0',user='moe',password='movies4me',dbname='IMDB',port=23306)
capt <- dbGetQuery(imcon$con,'SET NAMES utf8')

# values over 300M are likely bogus
budgets <- tbl(imcon,'movie_budgets') %>%
	filter(units=='$') %>%
	select(movie_id,amount) %>%
	filter(amount < 3e8)

role_info <- tbl(imcon,'role_type') %>%
	filter(role %in% c('actor','actress')) %>%
	select(role_id)

aggbudget <- tbl(imcon,'cast_info') %>%
	inner_join(role_info,by='role_id') %>%
	select(movie_id,person_id,role_id,nr_order) %>%
	inner_join(budgets,by='movie_id') %>%
	collect()

# fill with max
fix1 <- aggbudget %>%
	group_by(movie_id) %>%
	mutate(nr_order=pmin(nr_order,max(nr_order,na.rm=TRUE),na.rm=TRUE)) %>%
	ungroup()

# replace with 1
fix2 <- fix1 %>% 
	mutate(nr_order=pmin(nr_order,1,na.rm=TRUE)) 

# rescale
fix3 <- fix2 %>% 
	mutate(scale_order=1 / (nr_order)) %>%
	group_by(movie_id) %>%
	mutate(scale_order=scale_order / sum(scale_order,na.rm=FALSE)) %>%
  ungroup()

# multiply by budget, divide by 7, aggregate over people
bbyname <- fix3 %>% 
	mutate(salary=(1.0/7.0) * scale_order * amount) %>%
	group_by(person_id) %>%
	summarize(totsalary=sum(salary),
    totroles=n()) %>%
	ungroup()

# get name info
naminfo <- tbl(imcon,'name') %>%
	select(person_id,name,dob,gender) %>%
	filter(!is.null(dob)) %>%
	collect()

topcast <- bbyname %>% 
	inner_join(naminfo,by='person_id') %>%
	mutate(yob=as.numeric(gsub('^((19|20)\\d{2}).+','\\1',dob))) %>%
	filter(!is.na(yob)) %>%
  group_by(yob,gender) %>%
  arrange(desc(totsalary)) %>%
  summarize(totbudget=first(totsalary) / 1e6,
    topname=first(name),
    toproles=first(totroles)) %>%
  ungroup()


# I wish there were an easy way to do this via tidyr,
# and maybe there is...
topm <- topcast %>% 
  select(yob,topname,toproles,totbudget,gender) %>% 
	filter(gender=='m')
topf <- topcast %>% 
  select(yob,topname,toproles,totbudget,gender) %>% 
	filter(gender=='f')

topc <- inner_join(topm %>% select(-gender),topf %>% select(-gender),by='yob') %>%
	rename(actor_total_budget=totbudget.x,
    actor_name=topname.x,
    actor_roles=toproles.x,
    actress_total_budget=totbudget.y,
    actress_name=topname.y,
    actress_roles=toproles.y)

topc %>% 
	filter(yob > 1945,yob <= 1995) %>%
  kable()
```

-------------

## Dominant actors over time

Let us consider this budget calculation another way: compute the pseudo-salary for each
actor and actress as a function of production year of the film, 
then compute a 9 year rolling mean for each actor and actress, and 
plot the dominant players over time.

```{r starpower_III,cache=TRUE,eval=TRUE,tidy=FALSE,results='asis'}
library(RMySQL)
library(dplyr)
library(knitr)
imcon <- src_mysql(host='0.0.0.0',user='moe',password='movies4me',dbname='IMDB',port=23306)
capt <- dbGetQuery(imcon$con,'SET NAMES utf8')

# values over 300M are likely bogus
budgets <- tbl(imcon,'movie_budgets') %>%
	filter(units=='$') %>%
	select(movie_id,amount) %>%
	filter(amount < 3e8)

role_info <- tbl(imcon,'role_type') %>%
	filter(role %in% c('actor','actress')) %>%
	select(role_id)

movie_yr <- tbl(imcon,'title') %>%
	select(movie_id,production_year) %>%
	rename(year=production_year) 

aggbudget <- tbl(imcon,'cast_info') %>%
	inner_join(role_info,by='role_id') %>%
	select(movie_id,person_id,role_id,nr_order) %>%
	inner_join(budgets,by='movie_id') %>%
  inner_join(movie_yr,by='movie_id') %>%
	collect()

# fill with max
fix1 <- aggbudget %>%
	group_by(movie_id) %>%
	mutate(nr_order=pmin(nr_order,max(nr_order,na.rm=TRUE),na.rm=TRUE)) %>%
	ungroup()

# replace with 1
fix2 <- fix1 %>% 
	mutate(nr_order=pmin(nr_order,1,na.rm=TRUE)) 

# rescale
fix3 <- fix2 %>% 
	mutate(scale_order=1 / (nr_order)) %>%
	group_by(movie_id) %>%
	mutate(scale_order=scale_order / sum(scale_order,na.rm=FALSE)) %>%
  ungroup()

# multiply by budget, divide by 7, aggregate over people, years, summing
bbynameyr <- fix3 %>% 
	mutate(salary=(1.0/7.0) * scale_order * amount) %>%
	group_by(person_id,year) %>%
	summarize(totsalary=sum(salary),
    totroles=n()) %>%
	ungroup()

# this is godawful slow and terrible: compute salary on a rolling window
pyrs <- 1945:2016
rollsum <- function(yrnum,tots,window=9) {
  yrv <- rep(0,length(pyrs))
	isin <- pyrs %in% yrnum
  yrv[isin] <- tots[yrnum %in% pyrs]
  cyrv <- cumsum(yrv)
  rv <- data.frame(yr=pyrs[window:length(pyrs)],dsum=diff(cyrv,lag=window-1))
  rv[rv$dsum > 0,]
}

aggbyyr <- bbynameyr %>%
  filter(totsalary > 1000) %>%
	arrange(year) %>%
	group_by(person_id) %>%
	do(rollsum(.$year,.$totsalary)) %>%
  ungroup() %>%
	filter(dsum > 0)
```
```{r starpower_III_picktop,cache=TRUE,eval=TRUE,tidy=FALSE,results='asis'}
Kval <- 1

# select anyone who has been in the top K in any year, by gender
topK <- aggbyyr %>% 
	inner_join(naminfo %>% select(person_id,gender),by='person_id') %>%
  arrange(desc(dsum)) %>%
	group_by(yr,gender) %>% 
  summarize(cutoff=nth(dsum,Kval)) %>%
  ungroup()

intopK <- aggbyyr %>% 
	inner_join(naminfo %>% select(person_id,gender),by='person_id') %>%
  inner_join(topK,by=c('yr','gender')) %>%
  filter(dsum >= cutoff) %>%
	distinct(person_id)

# now get the top 
topdogs <- aggbyyr %>%
	inner_join(naminfo %>% select(person_id,name,dob,gender),by='person_id') %>%
  inner_join(intopK %>% select(person_id),by='person_id')

topdogs %>% 
	group_by(yr,gender) %>%
	arrange(desc(dsum)) %>%
	summarize(tname=first(name)) %>%
  ungroup() %>%
  rename(name=tname) %>%
  kable()
```
```{r starpower_III_plot,cache=TRUE,eval=TRUE,tidy=FALSE}
library(ggplot2)

ph <- ggplot(topdogs,aes(x=yr,y=dsum,group=name,color=name,label=name)) +
  geom_line() + 
  geom_text(check_overlap=TRUE) +
	labs(x='year',y='rolling quasi-salary ($)') +
  facet_grid(. ~ gender) + 
  scale_y_log10() + 
  guides(col=FALSE)

print(ph)
```


-------------

## Underrated movies

Here we use `dplyr` to connect to the database, and create a list of
the best 'underrated' movies of each year. The criteria are:

1. More than 500 ratings, but
1. fewer than 2000 ratings, and 
1. ranked via mean rating divided by standard deviation.


```{r sortem,cache=TRUE,eval=TRUE,tidy=FALSE}
library(RMySQL)
library(dplyr)
library(knitr)
imcon <- src_mysql(host='0.0.0.0',user='moe',password='movies4me',dbname='IMDB',port=23306)
dbGetQuery(imcon$con,'SET NAMES utf8')

tbl(imcon,'movie_votes') %>%
  filter(votes >= 500,votes <= 2000) %>% 
  inner_join(tbl(imcon,'movie_runtime') %>% 
    filter(runtime > 60) %>% 
    select(movie_id),by='movie_id') %>%
  mutate(vote_sr=vote_mean/vote_sd) %>%
  select(movie_id,vote_sr) %>%
  left_join(tbl(imcon,'title') %>% select(movie_id,title,production_year),by='movie_id') %>%
  arrange(desc(vote_sr)) %>%
  collect() %>% 
  group_by(production_year) %>%
  summarize(title=first(title),
    sr=max(vote_sr)) %>%
  ungroup() %>%
  arrange(production_year) %>%
  kable()
```



